---
title: "extract_data_Harz_2021"
format: html
---

### Harz 2020-2021

Script to extract the spectral values (bands and indices) and their temporal differences of two sentinel images (present, past) to the reference points with class damage and not damage and with damage type information.

```{r}
renv::restore
library(here)
library(terra)
library(data.table)
library(dplyr)
library(stringr)
source(local=TRUE, "functions.R")
```

### Load data

#### Load raster data

Load images from the tiles UNC and UNB for the years past and present.

```{r}
rpast <- rast(im2list("D:\\wsf-sat\\data\\validation\\Harz_2021\\Merge_UNC_UPC_2020_GTiff_shift_TC.tif"))
rpresent <- rast(im2list("D:\\wsf-sat\\data\\validation\\Harz_2021\\Merge_UNC_UPC_2021_GTiff_shift_TC.tif"))
names(rpast) <- paste(names(rpast), '_past', sep="")
names(rpresent) <- paste(names(rpresent), '_present', sep="")
```

#### Load geometry data

```{r}
# Geometries
damage_polygons <- vect("D:\\wsf-sat\\data\\validation\\Harz_2021\\Harz_Freiflaeche_stehendesTotholz_2021.gpkg", layer="filter025")
no_damage_points <- vect(here("data", "Referenz_NichtSchaden_randompoints_Harz_2021.gpkg"))

damage_polygons_table <- data.table(as.data.frame(damage_polygons))
no_damage_points_table <- data.table(as.data.frame(no_damage_points))
```

#### Prepare geometry data

Select only necessary columns and adjust values

Damage classes: 1 -\> stehendes Totholz, 2 -\> lliegendes Totholz, 3-\> Freifläche

```{r}
damage_polygons_table <- damage_polygons_table %>% 
                            rename(year = Jahr, damage_type = Typ_wsfsat) %>% 
                            select("year", "damage_type")  %>% 
                            mutate(damage_class = 1)
```

#### Calculate difference bands

Calculate and save difference raster present - past

```{r}
rdiff <- rpresent - rpast
names(rdiff) <- str_replace(names(rdiff), "_present", "_diff")

writeRaster(rdiff, filename=here("output", "Harz_2021_bands.tif"), filetype="GTiff", datatype='INT4S', overwrite=TRUE)  
```

#### Calculate indices and differences

Define bands and calculate indices

```{r}
# List with rasters
indices_past <- calculate_indices(rpast)
indices_present <- calculate_indices(rpresent)

names(indices_past) <- paste(names(indices_past), '_past', sep="")
names(indices_present) <- paste(names(indices_present), '_present', sep="")

# Multiband rasters
rindices_past <- rast(indices_past)
rindices_present <- rast(indices_present)
```

Calculate differences

```{r}
rindices_diff <- rindices_present - rindices_past
names(rindices_diff) <- str_replace(names(rindices_diff), "_present", "_diff")

writeRaster(rindices_past, filename=here("output", "Harz_2021_indices_past.tif"), filetype="GTiff", datatype='INT4S', overwrite=TRUE)  
writeRaster(rindices_present, filename=here("output", "Harz_2021_indices_present.tif"), filetype="GTiff", datatype='INT4S', overwrite=TRUE)  
writeRaster(rindices_diff, filename=here("output", "Harz_2021_indices_diff.tif"), filetype="GTiff", datatype='INT4S', overwrite=TRUE)  

```

#### Prepare extract raster

Join raster with bands values and raster with indices values

```{r}
rextract <- c(rpast, rpresent, rdiff, rindices_past, rindices_present, rindices_diff)
rm(rpast, rpresent, rdiff, rindices_past, rindices_present, rindices_diff)
```

#### Extract data to points

Extract image values to damage polygons and to no damage points

```{r}
extract_values_damage <- extract(rextract, damage_polygons) 
extract_values_no_damage <- extract(rextract, no_damage_points) 
```

Damage data: Merge table with the extracted values with table with attributes damage type, damage class and year

```{r}
damage_polygons_table$ID <- rownames(damage_polygons_table) # write rowid in column
extract_values_damage <- merge(extract_values_damage, damage_polygons_table)
```

No damage data: Add attributes damage type, damage class and year to the no damage table and create unique id that are not given in the damage table

```{r}
extract_values_no_damage$year <- 2021 
extract_values_no_damage$damage_type <- NA 
extract_values_no_damage$damage_class <- 0  
extract_values_no_damage$ID <- extract_values_no_damage$ID + max(extract_values_damage$ID)

```

Bind both tables: damage and no damage, create unique ID and save data

```{r}
extract_values_all <- rbind(extract_values_damage, extract_values_no_damage) 

d1 <- fread(here("output", "reference_Friedericke_2018_clean.csv"), dec = ",")
extract_values_all$ID <- extract_values_all$ID + max(d1$ID)

write.csv2(extract_values_all, here("output", "reference_Harz_2021.csv"), row.names=FALSE) 
```
