---
title: "mean_sd_analyse_detection"
format: html
---

Choose the best index for damage detection. Input spectral values from S2-images of the current and past year and the difference of these values + reference data as points with class damage and no damage. For each variable (index, band) two calculations are performed: the difference between the mean of the population damage and the population no damage and the mean value of the standard deviation of both populations. Best case scenario is a big difference between both classes and a small standard deviation in the classes. Results combining these two metrics as plot and as ratio.

From spectral values and indices from S2-images

```{r}
library(caret)
library(here)
library(data.table)
library(dplyr)
library(ggplot2)
```

Function definition:

-   Function to normalize values

-   Function to calculate the difference between the means of two populations

-   Function to calculate the mean value of the standard deviation of two populations

```{r}
min_max_norm <- function(x) {
    (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
  }

mean_diff <- function(value_damage, value_nodamage){
  mean1 <- lapply(value_damage, mean)
  mean2 <- lapply(value_nodamage, mean)
  diff <- mapply('-', mean1, mean2, SIMPLIFY=FALSE)
  abs_diff <- lapply(diff, abs)
  return(data.table(t(as.data.frame(abs_diff))))
}

sd_mean <- function(value_damage, value_nodamage){
  sd1 <- lapply(value_damage, sd)
  sd2 <- lapply(value_nodamage, sd)
  sd <- as.data.frame(rbind(unlist(sd1),unlist(sd2)))
  mean <- data.table(as.data.frame(colMeans(sd)))

  return(mean)
}
```

Read data: extracted spectral and index values to points. Data from Friedericke 2018 (mainly Freifläche) and from Harz 2021 (mainly stehendes Totholz)

```{r}
d1 <- fread(here("output", "reference_Friedericke_2018_clean.csv"), dec = ",")
d2 <- fread(here("output", "reference_Harz_2021_clean.csv"), dec = ",")
d <- rbind(d1,d2)

rm(d1,d2)
#d <- d %>% mutate_at("damage_class", as.factor)

# only columns diff values
d_diff <- d[ , c(grepl("diff" , names( d ))), with=FALSE ] 
```

Standardization of the spectral and index values

```{r}
d_norm <- as.data.frame((lapply(d_diff, min_max_norm))) 
d_norm <- cbind(d_norm, d[,"damage_class"]) # bind damage class column again
d_norm <- data.table(na.omit(d_norm))
```

Create two groups: damage and no damage and keep all columns except damage class because not needed from now on

```{r}
pop_nodam <- d_norm[d_norm$damage_class==0, -c("damage_class")] 
pop_dam <- d_norm[d_norm$damage_class==1, -c("damage_class")]
```

Apply functions: difference between population means and mean standard deviation of populations

```{r}
datamean <- mean_diff(pop_dam, pop_nodam)
datasd <- sd_mean(pop_dam, pop_nodam)
```

Create data frame

```{r}
names <- names(pop_dam)
dataplot <- cbind(names, datamean, datasd)
names(dataplot) <- c("var", "diff", "sd")
dataplot$ratio <- dataplot$diff / dataplot$sd
```

Plot

```{r}
ggplot(dataplot,aes(x=diff,y=sd,col=var,label=var))+geom_point()+geom_text(hjust=0, vjust=0)

```

```{r}
ggplot(dataplot, aes(x = reorder(var, -ratio), y = ratio, fill=var)) + geom_col() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("variable")
```
